---
title: "tidy_data_exercises_R4DS"
author: "Noelí D'Alessio"
date: "5/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Exercise 12.2.2

```{r}
library(tidyverse)
library(ggplot2)
```

2. Compute the rate for table2, and table4a + table4b. You will need to perform four operations:

Extract the number of TB cases per country per year.
Extract the matching population per country per year.
Divide cases by population, and multiply by 10000.
Store back in the appropriate place.

```{r}
table2_cases <- 
    filter(table2, type == "cases") %>% 
    rename(cases = count) %>% 
    arrange(country, year)

table2_population <- filter(table2, type == "population") %>% 
    rename(population = count) %>% 
    arrange(country, year)

t2_cases_per_cap <- 
    tibble(
    year = table2_cases$year,
    country = table2_cases$country,
    cases = table2_cases$cases,
    population = table2_population$population
          ) %>% 
    mutate(cases_per_cap = (cases / population) * 1000)
    
t2_cases_per_cap <- t2_cases_per_cap %>% 
    mutate(type = "cases_per_cap") %>% 
    rename(count = cases_per_cap)

view(t2_cases_per_cap)

t2<- bind_rows(table2, t2_cases_per_cap) %>% 
    arrange(country, year,type, count)
view(t2)
```


(table4a contains cases and table4b contains population)

```{r}
table4c <- 
    tibble(
      country = table4a$country,
      cases_cap_1999 = table4a[["1999"]] / table4b[["1999"]] * 1000,
      cases_cap_2000 = table4a[["2000"]] / table4b[["2000"]] * 1000
    )
```

Recreate the plot showing change in cases over time using table2 instead of table1. What do you need to do first?

```{r}
table2 %>% 
    filter(type == "cases") %>% 
    ggplot(aes(year, count)) +
    geom_line(aes(group = country), color= "grey50") +
    geom_point(aes(color = country)) +
    scale_x_continuous(breaks = unique(table2$year)) +
    ylab("cases")
```


# Pivoting

Why are pivot_longer() and pivot_wider() not perfectly symmetrical?
Carefully consider the following example:

stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17)
)
stocks %>% 
  pivot_wider(names_from = year, values_from = return) %>% 
  pivot_longer(`2015`:`2016`, names_to = "year", values_to = "return")
  
(Hint: look at the variable types and think about column names.)

pivot_longer() has a names_ptype argument, e.g. names_ptype = list(year = double()). What does it do?



