---
title: "Factors"
author: "Noelí D'Alessio"
date: "6/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Factors

```{r}
x1 <- c("Dec", "Jan", "Apr")

month_levels <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Ago", "Sep", "Oct", "Nov", "Dec")

y1 <- factor(x1, levels = month_levels)
y1

sort(y1)
```

```{r}
x2 <- c("Dec", "Jam", "Apr")
y2 <- parse_factor(x2, levels = month_levels)
```

# foorcats::gss_cat - General Social Survey

```{r}
gss_cat
```

### 15.3.1 Exercise

1. Explore the distribution of rincome (reported income). What makes the default bar chart hard to understand? How could you improve the plot?

```{r}
ggplot(gss_cat, aes(rincome)) +
    geom_bar() +
    scale_x_discrete(drop = FALSE) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

# the problem was the labels overlapping
```


Also, there are some ways to improve this bar chart:
-Removing "not applicable"
-Renaming Lt $ 1000 as Less than $1000
-Using colo to distinguish non-response categories (no answer, dk, refused)
-Adding meaningful labels
-Formating numbers with commas

```{r}
gss_cat %>% 
  filter(!rincome %in% c("Not applicable")) %>% 
  mutate(rincome = fct_recode(rincome, "Less than $1000" = "Lt $1000")) %>% 
  mutate(rincome_na = rincome %in% c("Refused", "Don't know", "No answer")) %>% 
  ggplot(aes(x = rincome, fill = rincome_na)) +
  geom_bar() +
  coord_flip() +
  scale_y_continuous("Number of Respondents", labels = scales::comma) +
  scale_x_discrete("Respondent's Income") +
  scale_fill_manual(values = c("FALSE" = "black", "TRUE" = "grey")) +
  theme(legend.position = "None")
```

2. What is the most common relig in this survey? What’s the most common partyid?

```{r}
gss_cat %>% 
  count(relig) %>% 
  arrange(desc(n))
```


```{r}
ggplot(gss_cat, aes(relig)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
gss_cat %>% 
  count(partyid) %>% 
  arrange(desc(n)) 
```

```{r}
ggplot(gss_cat, aes(partyid)) +
    geom_bar() +
    coord_flip()
```


3. Which relig does denom (denomination) apply to? How can you find out with a table? How can you find out with a visualisation?

```{r}
levels(gss_cat$denom)
```

### 15.4.1 Exercises

1. There are some suspiciously high numbers in tvhours. Is the mean a good summary?


```{r}
summary(gss_cat[["tvhours"]])
```

```{r}
gss_cat %>% 
  filter(!is.na(tvhours)) %>% 
  ggplot(aes(x = tvhours)) +
  geom_histogram(binwidth = 1)
```

2. For each factor in gss_cat identify whether the order of the levels is arbitrary or principled.

```{r}
keep(gss_cat, is.factor) %>% names()
```

This 6 are categorical variables. 

- In marital we could say there is a logical order:

```{r}
levels(gss_cat[["marital"]])
```
```{r}
gss_cat %>% 
  ggplot(aes(x = marital)) +
  geom_bar()
```

- In the case of the variable race, each category seems to be ordered accordig the number of observations:

```{r}
levels(gss_cat[["race"]])
```
```{r}
gss_cat %>% 
  ggplot(aes(x = race)) +
  geom_bar() +
  scale_x_discrete(drop = FALSE)
```

- In this case, the varibles have a decreasing order except fot the missing value's variables, which 3 of them are at the beggining and 1 at the end.This missing income level categories should be all together.

```{r}
levels(gss_cat[["rincome"]])
```

- There is a logic in this order, it works as a continumm from "Strong republican" to "Strong democrat".

```{r}
levels(gss_cat[["partyid"]])
```

- In the case of the variable religthe order of the levels is arbitrary.

```{r}
levels(gss_cat[["relig"]])
```
```{r}
ggplot(gss_cat, aes(x = relig)) +
  geom_bar() +
  coord_flip()
```

- denom

```{r}
levels(gss_cat[["denom"]])
```

- Denom doesn't seem to have a special order either.

```{r}
ggplot(gss_cat, aes(x = denom)) +
  geom_bar() +
  coord_flip()
```

3. Why did moving “Not applicable” to the front of the levels move it to the bottom of the plot?

Because that gives the level “Not applicable” an integer value of 1.

#### 15.5.1 Exercises

1. How have the proportions of people identifying as Democrat, Republican, and Independent changed over time?

```{r}
gss_cat %>% 
  mutate(
    partyid = 
        fct_collapse(partyid,
        other = c("No answer", "Don't know", "Other party"),
        rep = c("Strong republican", "Not str republican"),
        ind = c("Ind,near rep", "Independent", "Ind,near dem"),
        dem = c("Not str democrat", "Strong democrat")
        )
    ) %>% 
  count(year, partyid) %>% 
  group_by(year) %>% 
  mutate(p = n / sum(n)) %>% 
  ggplot(aes(x = year, y = p, colour = fct_reorder2(partyid, year, p))) +
  geom_point() +
  geom_line() +
  labs(colour = "Party ID")
```


2. How could you collapse rincome into a small set of categories?

```{r}
levels(gss_cat[["rincome"]])
```


```{r}
library("stringr")
```
```{r}
gss_cat %>% 
    mutate(
      rincome =
        fct_collapse(
          rincome, 
          `Unknown` = c("No answer", "Don't know", "Refused", "Not applicable"),
          `Lt $5000` = c("Lt $1000", 
          str_c("$", c("1000", "3000", "4000"), " to ", c("2999", "3999", "4999"))),
          `$5000 to 10000` = 
            str_c("$", c("5000", "6000", "7000", "8000"), " to ",
            c("5999", "6999", "7999", "9999"))
        )
      ) %>% 
  ggplot(aes(x = rincome)) +
  geom_bar() +
  coord_flip()
```

