---
title: "Exploratory_data_analysis_R4DS"
author: "Noelí D'Alessio"
date: "5/4/2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Exploratory Data Analysis

```{r}
library("tidyverse")
library("viridis")
library("nycflights13")
library("ggbeeswarm")
library("lvplot")
library("ggstance")
```

### 1. Explore the distribution of each of the x, y, and z variables in diamonds. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.

x length in mm (0-10.74)
y width in mm (0-58.9)
z depth in mm (0-31.8)
```{r}
summary(select(diamonds, x, y, z))
```


```{r}
ggplot(data = diamonds) +
    geom_histogram(mapping = aes(x = x), binwidth = 0.01)
```

```{r}
ggplot(data = diamonds) +
    geom_histogram(mapping = aes(x = y), binwidth = 0.01)
```

```{r}
ggplot(data = diamonds) +
    geom_histogram(mapping = aes(x = z), binwidth = 0.01)
```

1. X is larger than Y, and Y is larger than Z. 
2. There are outliers in the 3 variables. 
3. They have a positive, right asimetry.
    this makes sense, there should be more "small diamonds" than larger. And can't be diamonds with 0 mm.
4. They have a multimodal distribution. 
    Related to the caract? Maybe there are some "cuts"?

Typical values of X go from 4.710 to 6.540.
Typical values of Y go from de 4.720 to 6.540.
Typical values of Z go from 2.910 to 4.040.
Es decir, los valores típicos de X e Y are larger than Z's values.

Otliers values: 0 and some are large values for X, Y, and Z. 

```{r}
filter(diamonds, x == 0 | y == 0 | z == 0)
```

0 values, missing values?

```{r}
diamonds %>% 
    arrange(desc(y)) %>% 
    head(20)
```
```{r}
diamonds %>% 
    arrange(desc(z)) %>% 
    head(20)
```
Also did the procedure, arranging descending order by using the X value but X didn't have any unusual value.

Others outliers:
Y= 58.90?, 31.80?
Z= 31.80?

To check the consistency of this values with other variables, we carry a scatter plots of each combination X, Y, Z. This shows outliers more clearly.

```{r}
ggplot(diamonds, aes(x = x, y = y)) +
  geom_point()
```
```{r}
ggplot(diamonds, aes(x = x, y = z)) +
    geom_point()
```

```{r}
ggplot(diamonds, aes(x = z, y = y))+
  geom_point()
```

To filter the outliers:
```{r}
filter(diamonds, x > 0, x < 10) %>% 
  ggplot() +
  geom_histogram(mapping = aes(x = x), binwidth = 0.01) +
  scale_x_continuous(breaks = 1:10)
```
```{r}
filter(diamonds, y > 0, y < 10) %>% 
    ggplot()+
    geom_histogram(mapping = aes(x = y), binwidth = 0.01) +
    scale_x_continuous(breaks = 1:10)
```

```{r}
filter(diamonds, z > 0, z < 10) %>% 
    ggplot()+
    geom_histogram(mapping = aes(x = z), binwidth = 0.01) +
    scale_x_continuous(breaks = 1:10)
```

### 2. Explore the distribution of price. Do you discover anything unusual or surprising? (Hint: Carefully think about the binwidth and make sure you try a wide range of values.)

```{r}
summary(select(diamonds, price))
```

```{r}
ggplot(data = diamonds) +
    geom_histogram(mapping = aes(x = price), binwidth = 100, center = 0)
```

Not much information in the previous graphic.

```{r}
ggplot(filter(diamonds, price < 2500), aes(x = price)) +
    geom_histogram(binwidth = 10, center = 0)
```
-Hay varios picos en la data pero no sé a qué corresponden.
-Hay un pico en la distribución en el precio de $750.
-No hay diamantes con un precio de $1500 (entre $1450 y $1550).

######Distribución del último dígito

```{r}
diamonds %>% 
    mutate(ending = price %% 10) %>% 
    ggplot(aes(x = ending)) +
    geom_histogram(binwidth = 1, center = 0)
```


```{r}
diamonds %>% 
    mutate(ending = price %% 100) %>% 
    ggplot(aes(x = ending)) +
    geom_histogram(binwidth = 1)
```

```{r}
diamonds %>% 
    mutate(ending = price %% 1000) %>% 
    filter(ending >= 500, ending <= 800) %>% 
  ggplot(aes(x = ending)) +
  geom_histogram(binwidth = 1)
```

### How many diamonds are 0.99 carat? How many are 1 carat? What do you think is the cause of the difference?

```{r}
diamonds %>% 
    filter(carat >= 0.99, carat <= 1) %>% 
    count(carat)
```

One reason could be the round up the carat to 1. 

```{r}
diamonds %>% 
    filter(carat >= 0.9, carat <= 1.1) %>% 
    count(carat) %>% 
    print(n = Inf)
```
There are not many diamonds with a carat between 0.93 and 0.99.

### library("ggstance")

Instead of using coord_flip() to have horizontal boxplot we can use geom_boxploth from the library "ggstance":

```{r}
ggplot(data = mpg) +
    geom_boxploth(mapping = aes(y = reorder(class, hwy, FUN = median), x = hwy))
```

### With large data, instead of boxploth (which displays "outlying values") geom_lv drom the library ("lvplot"). 

```{r}
ggplot(diamonds, aes(x = cut, y = price)) +
  geom_lv()
```


```{r}
diamonds %>% 
    count(color, cut) %>% 
    group_by(cut) %>% 
    mutate(prop = n / sum(n)) %>% 
    ggplot(mapping = aes(x = color, y = cut)) +
    geom_tile(mapping = aes(fill = prop)) +
    scale_fill_viridis(limits = c(0, 1))
```

```{r}
flights %>% 
    group_by(month, dest) %>% 
    summarise(dep_delay = mean(dep_delay, na.rm = TRUE)) %>% 
    ggplot(aes(x = factor(month), y = dest, fill = dep_delay)) +
    geom_tile() +
    labs(x = "Month", y = "Destination", fill = "Average of Departure Delay")
    
```


To improve it we could remove the NA's. Also, sort destination by a meaninfulg quatinty. And better colors for the graphic. 

```{r}
flights %>% 
    group_by(month, dest) %>% 
    summarise(dep_delay = mean(dep_delay, na.rm = TRUE)) %>% 
    group_by(dest) %>% 
    filter(n() == 12) %>% 
    ungroup() %>% 
    mutate( dest = reorder(dest, dep_delay)) %>% 
    ggplot(aes(x = factor(month), y = dest, fill = dep_delay)) +
    geom_tile() +
    scale_fill_viridis() +
    labs(x = "Month", y = "Destnation", fill = "Departure Delay")
```
    
# Covariation with 2 continuous variables 

1. Instead of summarising the conditional distribution with a boxplot, you could use a frequency polygon. What do you need to consider when using cut_width() vs cut_number()? How does that impact a visualisation of the 2d distribution of carat and price?

cut_number: we specify how many bind we want. Numbers in each bin are calculated automatically.
```{r}
ggplot(data = diamonds, mapping = aes(color = cut_number(carat, 5), x = price)) +
  geom_freqpoly() +
  labs(x = "Price", y = "Count", color = "Carat")
```

cut_width: we specify the width of the bin. For example, we could choose 2, but there are not that many amount of diamonds with a carat larger than 2, so better smaller. 0.5 could be another option but it would create just too many bins (12), so we choose 1 as the width. 

```{r}
ggplot(data = diamonds, mapping = aes(color = cut_width(carat, 1, boundary = 0), x = price)) +
    geom_freqpoly() +
    labs(x = "Price", y = "Count", color = "Carat")
```

2. Visualise the distribution of carat, partitioned by price.

First, we will part the continuos variable of price into 10 bins:
```{r}
ggplot(data = diamonds, mapping = aes(x = cut_number(price, 10), y = carat)) +
    geom_boxplot(varwidth = TRUE) +
    coord_flip() +
    xlab("Price")
```

Now we are going to determine the width of the bins as $2000.

```{r}
ggplot(diamonds, aes(x = cut_width(price, 2000, boundary = 0), y = carat)) +
    geom_boxplot(varwidth = TRUE) +
    coord_flip() +
    xlab("Price")
```

3. How does the price distribution of very large diamonds compare to small diamonds. Is it as you expect, or does it surprise you?

4. Combine two of the techniques you’ve learned to visualize the combined distribution of cut, carat, and price.



