---
title: "Dates and Times exercises"
author: "Noelí D'Alessio"
date: "7/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("lubridate")
library("nycflights13")
```

16.2.4 Exercises

1. What happens if you parse a string that contains invalid dates?

```{r}
ymd(c("2010-10-10", "bananas")) 
```

It produces a NA value.

2.What does the tzone argument to today() do? Why is it important?

```{r}
today(tzone = "Africa/Banjul")
```
```{r}
now(tzone = "Africa/Banjul")
```

A lis of the different tzones: 

https://en.wikipedia.org/wiki/List_of_tz_database_time_zones

3. Use the appropriate lubridate function to parse each of the following dates:
```{r}
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14" # Dec 30, 2014
```

```{r}
mdy(d1)
ymd(d2)
dmy(d3)
mdy(d4)
mdy(d5)
```

16.3.4 Exercises

1. How does the distribution of flight times within a day change over the course of the year?

2. Compare dep_time, sched_dep_time and dep_delay. Are they consistent? Explain your findings.

3. Compare air_time with the duration between the departure and arrival. Explain your findings. (Hint: consider the location of the airport.)

4. How does the average delay time change over the course of a day? Should you use dep_time or sched_dep_time? Why?

5. On what day of the week should you leave if you want to minimise the chance of a delay?

5. What makes the distribution of diamonds$carat and flights$sched_dep_time similar?

6. Confirm my hypothesis that the early departures of flights in minutes 20-30 and 50-60 are caused by scheduled flights that leave early. Hint: create a binary variable that tells you whether or not a flight was delayed.
