---
title: "Dates and Times recap"
author: "Noelí D'Alessio"
date: "7/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(nycflights13)
```

### Creating date and date-times

```{r}
today()
```
```{r}
now()
```
Using lubridate's helpers:
```{r}
mdy("January 31st, 2019")
```
Also works with unquoted numbers
```{r}
ymd(20200721)
```

### Creating date/time from a string using yms_hms()

To create date-times we use ymd_hms():
```{r}
ymd_hms("2020_07_21 11-45-00")
```

### Creating date/time from individual components with make_date() and make_datetime():

```{r}
flights %>% 
    select(year, month, day, hour, minute) %>% 
    mutate(departure = make_datetime(year, month, day, hour, minute))
```

# Creating a function to transform date time

```{r}
make_datetime_100 <- function(year, month, day, time){
  make_datetime(year, month, day, time %/% 100, time %% 100)
}
```

Disclaimer: it would have been better to make dep_time by adding the dep_delay to the sched_dep_time. 

```{r}
flights_dt <- flights %>% 
  filter(!is.na(dep_time), !is.na(arr_time)) %>% 
  mutate(
    dep_time = make_datetime_100(year, month, day, dep_time),
    arr_time = make_datetime_100(year, month, day, arr_time),
    sched_dep_time = make_datetime_100(year, month, day, sched_dep_time),
    sched_arr_time = make_datetime_100(year, month, day, sched_arr_time)
  ) %>% 
  select(origin, dest, ends_with("delay"), ends_with("time"))

flights_dt
```

We can now visualise the distribution of departure time across the year:

```{r}
flights_dt %>% 
  ggplot(aes(dep_time)) +
  geom_freqpoly(binwidth = 86400) # 86400 seconds = 1 day
```

Or within a single day:

```{r}
flights_dt %>% 
  filter(dep_time < ymd(20130102)) %>% 
  ggplot(aes(dep_time)) + 
  geom_freqpoly(binwidth = 600) # 600 s = 10 minutes
```

Note that when you use date-times in a numeric context (like in a histogram), 1 means 1 second, so a binwidth of 86400 means one day. For dates, 1 means 1 day.

### Creating dates from other types

```{r}
as_datetime(today())
```
```{r}
as_date(now())
```

16.3 Date-time components

Useful functions: year(), month(), mday(), yday(0), wday(), hour(), minute(), second().
month(x, label = TRUE, abbr = FALSE)
wday(x, label = TRUE, abbr = FALSE)

```{r}
flights_dt %>% 
  mutate(wday = wday(dep_time, label = TRUE)) %>% 
  ggplot(aes(wday)) +
  geom_bar()
```


In the next plot we see that around flights departing around the minutes 20-30' and 50-60' have a lower average delay.
```{r}
flights_dt %>% 
  mutate(minute = minute(dep_time)) %>% 
  group_by(minute) %>% 
  summarise(
    avg_delay = mean(arr_delay, na.rm = TRUE),
    n = n()) %>% 
  ggplot(aes(minute, avg_delay)) +
  geom_line()
  
```

But, there doesn't seem to be a pattern in the schedule departure time and the avergae delay.
```{r}
sched_dep <- flights_dt %>% 
  mutate(minute = minute(sched_dep_time)) %>% 
  group_by(minute) %>% 
  summarise(
    avg_delay = mean(arr_delay, na.rm = TRUE),
    n = n())

ggplot(sched_dep, aes(minute, avg_delay)) +
  geom_line()
```

If we compare the distributions between schedule_departure and dep_time we see that there are more scheduled departures around 0' and 30', and the actual departure time around 30' and 60'. This might be a human biased, rounding the time.

```{r}
ggplot(sched_dep, aes(minute, n)) +
  geom_line()
```
```{r}
flights_dt %>% 
  mutate(minute = minute(dep_time)) %>% 
  group_by(minute) %>% 
  summarise(
    n = n()
  ) %>% 
  ggplot(aes(minute, n)) +
  geom_line()
```

### 16.3 Rounding:

```{r}
flights_dt %>% 
  count(week = floor_date(dep_time, "week")) %>% 
  ggplot(aes(week, n)) +
  geom_line()
```

### update():

```{r}
flights_dt %>% 
  mutate(dep_hour = update(dep_time, yday = 1)) %>% 
  ggplot(aes(dep_hour)) +
  geom_freqpoly(binwidth = 300) # 5 minutes
```


# Time spans


## Durantion

```{r}
noeli_age <- today() - ymd(19930305)
noeli_age
```

lubridate fx duration()
```{r}
as.duration(noeli_age)
```


## Periods

- There are some flights that in the table appear as they had arrived before than what they have left. This is because they are overnight flights. They arrived in the next day, but as we use the same information to construct this column, we have this mistake.

```{r}
flights_dt %>% 
  filter(arr_time < dep_time)
```

To fix it, we will add a day:

```{r}
flights_dt <- flights_dt %>% 
  mutate(
    overnight = arr_time < dep_time,
    arr_time = arr_time + days(overnight * 1),
    sched_arr_time = sched_arr_time + days(overnight * 1)
)
```

```{r}
flights_dt %>% 
  filter(overnight, arr_time < dep_time)
```

## Intervals

```{r}
dyears(1) / ddays(365)
```

```{r}
years(1) / days(1)
```

```{r}
next_year <- today() + years(1)
(today() %--% next_year) / ddays(1)
```

```{r}
(today() %--% next_year) %/% days(1)
```

